\protect Connecting Every Bit of Knowledge: \\ Wikipedia's First Link Network Unravelled

### 2015-12-30 writeup/2015-11-wikipedia-network.tex
edits to intro + method

\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}

@@

\affiliation{Department of Mathematics \& Statistics, 
    Computational Story Lab,  \\
    The University of Vermont, Burlington, VT 05401.}

@@

\affiliation{Department of Mathematics \& Statistics, 
    Computational Story Lab,  \\
    The University of Vermont, Burlington, VT 05401.}

\begin{abstract}
  \protect
Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic though, Wikipedia is a web of naturally emerging relationships.  
By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
Here we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
are related and organized.

We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
In a novel approach to uncover structure, we traverse every possible path through the network, 
measuring the accumulation of First Links, path lengths, basins, cycles, and even the influence a particular article exerts in shaping the 
network.
We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
culminating around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of how we connect and structure
an ever growing load of information.

@@

Wikipedia is a towering achievement of the modern era. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{stats} \cite{drugs}.
All this information does not lie dormant. 
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views in a single month
\cite{wiki_views}.
The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.

It's not surprising then to see Wikipedia as the object of many studies. 
Researchers have examined the cultural dynamics among editors
\cite{editors},
the accuracy of the content relative to traditional encyclopedias
\cite{accuracy1},
\cite{accuracy2},
the disciplines and topics covered 
\cite{coverage},
and bias against portions of the population
\cite{bias_women}.
Wikipedia's content has also proven a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{missing_entries},
cluster short text
\cite{clustering},
compute semantic relatedness
\cite{semantic_relatedness},
and disambiguate meaning \cite{disambiguating}.

While these many studies have dissected and fruitfully applied Wikipedia's content,
few have examined the connections among the many articles.
Beyond information, Wikipedia is a web 
relating ideas. 
Within the body text of Wikipedia articles, hyperlinks reference
other Wikipedia articles.
A hyperlink from one Wikipedia article to another is a way to relate the ideas
conveyed in the two articles 
\cite{relevance}.
The notion that hyperlinks convey information about the content of a
page has proved successful in numerous domains from search engine algorithms 
such as PageRank 
\cite{pagerank} 
to topic classification
\cite{classifier}.
We treat a hyperlink as a mechanism to connect two ideas.

The authors of a Wikipedia article choose where and whether to include a 
hyperlink reference to another Wikipedia article
by including hyperlinks in the article's markup.
In describing a particular topic, the authors decide which other articles are most
relevant references.
For example, the authors of the "Train" article chose
"Rail Transport," "Cargo," and "Passengers" 
as pertinent articles to reference in describing trains.
The collection of hyperlinks emerges from the independent choice of each article's authors. 
The first hyperlink to another Wikipedia article marks the earliest moment in the body text where the authors
choose to connect the information discussed to that of another article.
By following the first hyperlink to another article in the English edition of
Wikipedia, we connect the ideas in one article to those in another, ultimately forming a directed network: 
{\it Wikipedia's First Link Network} (FLN).

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
  \caption{
    \textbf{First Link Path For "Train"}
    We follow the First Link to another Wikipedia article
    in the main body of the article---the area
    inside the green rectangle, which excludes 
    side bar elements, the navigation bar and title. 
    In this example, the First Link to another Wikipedia article is "Rail Transport." 
    We can again select the First Link on the "Rail Transport"
    article, repeating the process to 
    form a path of First Links.
  }
  \label{fig:Train First Links}
\end{figure*}

What information could the First Links possibly reveal?
Inspired by the claim that the majority of First Links lead to the 
"Philosophy" article---popularized by an xkcd comic and subsequently
discussed in blog posts 
\cite{mat_blog}
\cite{Ilmari_first_links}
\cite{xkcd} 
---we holistically study 
Wikipedia's FLN as a map relating areas of human knowledge. 
The map is a hierarchical taxonomy where "Train" links to a parent node, 
"Rail Transport," while many child nodes, "Steel" and "Horsepower," link up to 
"Train." Unlike previous taxonomies
created by individuals
\cite{locke}
\cite{descartes}
\cite{aristotle} 
or a select group of individuals 
\cite{hist_thesaurus}, 
the organization of ideas in the FLN 
does not emerge from a concerted or a centralized effort. 
Rather each link in the FLN emerges from the independent choice of an article's authors.
Collectively, the FLN is a wealth of relations among inventions, places,
figures, objects, and events across space and time.

Our goal is to gain insight into how the information on Wikipedia is organized and related
by studying the structure of the FLN.
In a novel approach, we develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of First Links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, indegree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

One essential measure of a directed network structure is the degree distribution 
\cite{newman}. 
Degree distribution has been used to study many phenomena from disease outbreak 
\cite{disease} 
to the dynamics of social networks 
\cite{social_nets}.
The degree distribution in the FLN describes how many First Links point to a particular article. 
We measure the indegree in the FLN to guage the number of articles referencing a particular article.
A higher indegree indicates more authors selected the particular article
as a relevant reference. On the other hand, articles with zero 
indegree have no references---they are outer leaves in the FLN. 
Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
We can rank articles by indegree to see which articles are most and least referenced in the FLN. 
Additionally, we can assess the degree distribution to see whether authors tend to reference all articles
equally or only a few, as observed in many social networks \cite{social_nets}.

The degree distribution captures one step in the FLN, measuring   
the destination of a given link. 
Indegree measures only the particular set of First Links between two articles, 
rather than the richer dynamics of how many articles fit together---how links 
create a flow through the FLN.
One way to expand the analysis beyond a single step is to map the full path of 
an article's First Link. 
Each article is both a destination and a link to another article, similar to an element 
in a linked list data structure. 
By following the chain of links we can construct a path through the FLN with each article 
leading to the next via the First Link. 
For example "Train" has a First Link to "Conveyance of Passengers and Goods," which is itself
an article with a First Link to "Goods" and so on. The path starting at "Train" 
is the collection of articles obtained by following the First Links starting at "Train"
(see figure~\ref{fig:Train First Links}).
The path contains "Goods," "Economics," "Social Science," and other articles related to 
"Train," but not directly referenced by a First Link.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
  \caption{
    \textbf{Traversal Visit Algorithm on a sample network}
     The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A. The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix:
     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
  }
  \label{fig:Traversal Visits}
\end{figure*}

Applying a similar concept with a 
different starting article, we form 
a path through the FLN moving from one article to the next.
A path in the FLN connects a group of articles allowing us to measure how two articles relate,
even if the articles are not directly linked. 
The collection of articles is path-connected conveying a flow of concepts from one article to another. 
The notion of paths allows us to map not only which articles link directly to "Philosophy" 
for example, but also which ideas are eventually anchored in "Philosophy" even if the articles
are not directly linked.
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Starting at each article, we construct a path through the FLN for each article, 
ultimately mapping how the First Links flow through the FLN.

@@

Our goal is to map the flow of First Links by following the paths of First Links 
starting at each article. 
Since a path reveals how the ideas in one article may eventually be 
anchored in another, the first metric we develop quantifies the accumulation of First Links 
within the FLN. 
The algorithm begins by selecting an article then traversing the path formed
by following the First Links. Each time a First Link references an article, we increment a count
associated with the article. 
We repeat until the First Link is repeated or invalid (outside the FLN). 
We select a second article and repeat the process until we have 
constructed a path for each article in the network. We call the resulting count for each article the number of {\it traversal visits}. The number of traversal visits of an article 
measures the number of references flowing to the ideas in the article.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
  \caption{
    \textbf{Traversal Funnel Algorithm on a sample network}
  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path. We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row. For example
  the number of traversal funnels for article $E$ is 
  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
  \label{fig:Traversal Funnels}

\end{figure*}

We can map the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
(see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
first column in the traversal vists matrix. 
An entry of $1$ indicates the path contains a given article and 
$0$ indicates the path doesn't. The columns identify the starting article of the path. 
For example, the second column corresponds to the path starting at article B. 
To compute the number traversal visits for article A we sum the corresponding row in our matrix.
The total is the number of paths containing article A. 
Analogous to the flow 
in a river network, traversal visits measure the accumulation of water near point A. 
By constructing a traversal visits matrix, we can similarly compute the number of traversal 
visits by summing along the corresponding row for our node. The traversal visits matrix for 
Wikipedia consists 121 million entries to account for each path through the FLN. 
We measure the accumulation of First Link references by summing the corresponding row for the 
article of interest.

In addition to accumulation, the traversal visits algorithm creates paths 
which uncover depth in the FLN. 
By measuring the number of First Links between
two articles, we get an additional piece of information we call 
{\it path length.} We can compute the length a path by summing 
along a column in our traversal visits matrix. 
The path length describes how closely related ideas are, gauging the FLN's depth.
The greater the number of First Links separating two articles, the greater the number of ideas needed to relate the ideas. Although, 
"Train" is related to "Economics," there are several more specific articles
bridging the connection:
"Train" is more specifically related to transportation, whose object
is often goods. Goods are one of the fundamental objects of study in Economics.
Described in links, this relationship is captured by a relatedness of $4$ 
First Links ultimately connecting "Train" to "Economics." 
We can examine path length as a relative measure of how closely connected ideas
may be, but also as an aggregate measure describing the overall separation or 
connectedness in the FLN.
We can characterize whether ideas a closely connected by a handful of links, 
many, or whether a more nuanced organization exists.

One possible path through the FLN is a {\it cycle} or a group of articles 
linking to one another inside a loop. In our sample network 
(see figure~\ref{fig:Traversal Visits}) 
a cycle exists among nodes A, B, and C. 
Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
By recording the history of the articles along a path in the FLN, we can identify
the types of cycle structures within the FLN. Furthermore, since each article 
in the a cycle has an associated traversal visit count, we can rank cycles by 
the number of references directed towards a cycle. 
We can also form and rank {\it basins} in the network by identifying groups of 
path-connected articles, not necessarily forming a perfect cycle.
A basin connects a group of articles which ultimately reference other articles
in the FLN. By identifying basins and cycles, we pinpoint groups of closely related articles.

While traversal visits measure accumulation, each article's First Link also 
influences the shape of the FLN. 
At a large point of accumulation, a single article's First Link 
can exert great influence over the shape of the FLN by directing many
references towards a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many First Links, we develop a second metric called {\it traversal funnels}.

To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle from an article exerting influence over where the First Links flow. 
An article with a large number of traversal funnels directs many references
towards a particular path. In our sample network 
(see figure~\ref{fig:Traversal Funnels}) article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
receipients of the flow---without exerting any influence themselves. 
While articles A, B, and C all have the same number of traversal visits, 
article C has 4 traversal funnels (A and B have none). By incrementing the count of 
First Link references on paths only up to a cycle, we can identifying which articles
exert the greatest influence over the FLN's structure.

By studying the FLN not only as collection of directly linked pairs of articles, but
as a flow of paths, 
we build a powerful arsenal of information with which to study the FLN. 
From accumulated references, cycles, basins to influence, we can measure how the many articles in Wikipedia are organized and related.

@@

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
  \caption{
    \textbf{Highest Ranking Articles by indegree}
  }
  We rank each article by the number of direct First Links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
  comprised of various facets and parts.
  \label{fig:ndegree list}
\end{figure}

We first identify the set of First Links directly referencing a particular article 
by studying the indegree of the FLN. 
A higher indegree suggests more authors found the ideas in the article 
pertinent to reference. 
We rank all 11 million articles by indegree to find 
the "United States" with 80,249 direct First Links as the most referenced
Wikipedia article. 
Other high-ranking articles
include foundational abstract concepts such as "village," "species,"; 
sports associations such as "American Football," "Association Football"; 
and developed nations such as "France," "Japan," "Russia," "Australia," and 
the "Netherlands." These high ranking articles are useful abstractions: nations
describe a collection of individuals, a common culture, language, or 
geographical proximity; sports teams describe an ever changing collection of 
sports players often associated with a cultural identity and a geographical 
region. Since abstractions such as nations and teams are inherently comprised
of many parts, authors reference the larger abstraction when describing 
a part. For example, in describing the geographical region of north America, 
the United States, Canada, or Mexico are natural references anchoring the 
geographical location to a larger national identity.

"Philosophy" and other philosophical concepts with many traversal visits
are not among the highest-ranking articles by indegree.
"Philosophy" has an indegree of 581, with direct First Links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy" (see traversal visits discussion), 
the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of First Links, as we argue in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears to be scale free with most articles having fewer than 9 
  direct First Links, while few hold most.
  \label{fig:degree distribution}
\end{figure}

The FLN's indegree exhibits a scale-free distribution where a few articles 
receive a most direct first references, while most articles receive few or none.
The average indegree for all 11 million articles is 3.6 direct First Links with a standard deviation of 89.5 links.
Only 4,826 articles have more than 100 direct First Links and $75\%$ of articles
have fewer than 9. 
When fit to a linear model on a log-log scale (log(rank) versus log(indegree)), 
the model's r value is $-0.98$ suggesting a strong log-log linear fit 
with a power law exponent of $-0.79$.

@@

Beyond indegree, our work expands the analysis of relations to understand 
how many more than two articles are organized and related. 
We develop the notion of a path to describe the structure of the FLN as a 
flow relating ideas. We traverse every possible path through the FLN, taking 
$232$ million steps along the way.

@@

We first seek to describe the depth of the FLN: how many links does a 
connection of ideas span? We gauge the FLN's depth via path length or the number of links traversed until a repeated or invalid link (one outside the FLN). 
We discover the longest path length is 365
corresponding to the yearly calendar of Orthodox Liturgics.
Each day's Liturgics links to the next day's. Curiously, on the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
(see discussion of cycles).
We also found similarly lengthy paths following the evolution of a place or topic through time: 
"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
The longest paths connecting ideas are organized temporally.

>>>>

\usepackage{hyperref}
\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}

@@

\affiliation{Department of Mathematics \& Statistics, 
    Computational Story Lab,  \\
    Vermont Complex Systems Center,
    Vermont Advanced Computing Core,
    The University of Vermont, Burlington, VT 05401.}

@@

\affiliation{Department of Mathematics \& Statistics, 
    Computational Story Lab,  \\
    Vermont Complex Systems Center,
    Vermont Advanced Computing Core,
    The University of Vermont, Burlington, VT 05401.}

\begin{abstract}
  \protect
Apples, oranges, and the most obscure Dylan song too---is every subject a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic though, Wikipedia is a web of naturally emerging relationships.  
By following the First Link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
Here we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
are related and organized.

We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
In a novel approach to uncover structure, we traverse every possible path through the network, 
measuring the accumulation of First Links, path lengths, basins, cycles, and even the influence a particular article exerts in shaping the 
network.
We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general 
culminating around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of how we connect and structure
an ever growing load of information.

@@

Wikipedia is a towering achievement of the modern era. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{drugs} 
\cite{stats}.
All this information does not lie dormant. 
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views and 26 million edits in a single month
\cite{wiki_edits}
\cite{wiki_views}.
The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.

It's not surprising then to see Wikipedia as the object of many studies. 
Researchers have examined the cultural dynamics among editors
\cite{editors},
the accuracy of the content relative to traditional encyclopedias
\cite{accuracy1}
\cite{accuracy2},
the topics covered 
\cite{coverage},
and bias against portions of the population
\cite{bias_women}.
Wikipedia's content has also proven a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{missing_entries},
cluster short text
\cite{clustering},
compute semantic relatedness
\cite{semantic_relatedness},
and disambiguate meaning \cite{disambiguating}.

While these many studies have dissected and fruitfully applied Wikipedia's content,
few have examined the connections among the many articles.
Beyond information, Wikipedia is a web 
relating ideas. 
Within the body text of Wikipedia articles, hyperlinks reference
other Wikipedia articles.
A hyperlink from one Wikipedia article to another
indicates a relationship between the two articles
\cite{relevance}.
The notion that hyperlinks convey information about the content of a
page has proved successful in numerous domains from search engine algorithms 
such as PageRank 
\cite{pagerank} 
to topic classification
\cite{classifier}.
We treat a hyperlink as a mechanism connecting two ideas.

The authors of a Wikipedia article choose where and whether to include a 
reference to another Wikipedia article
in the HTML markup.
In describing a particular topic, the authors decide which other articles are most
relevant references.
For example, the authors of the 
\href{https://en.wikipedia.org/wiki/Train}{"Train"}
article chose
"Rail Transport," "Cargo," and "Passengers" 
as pertinent articles to reference in describing trains
\cite{wiki_train}
.
The collection of hyperlinks emerges from the independent choice of each article's authors. 
The first hyperlink to another Wikipedia article marks the earliest moment in the body text where the authors
choose to connect the information discussed to that of another article.
By following the first hyperlink to another article in the English edition of
Wikipedia, we connect the ideas in one article to those in another, ultimately forming a directed network: 
{\it Wikipedia's First Link Network} (FLN).

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
  \caption{
    \textbf{First Link Path For "Train".}
    We follow the First Link to another Wikipedia article
    in the main body of the article---the area
    inside the green rectangle, which excludes 
    side bar elements, the navigation bar and title. 
    In this example, the First Link to another Wikipedia article is "Rail Transport." 
    We can again select the First Link on the "Rail Transport"
    article, repeating the process to 
    form a path of First Links.
  }
  \label{fig:Train First Links}
\end{figure*}

What information could the First Links possibly reveal?
Inspired by the claim that the majority of First Links lead to the 
"Philosophy" article---popularized by an 
\href{https://xkcd.com/903/}{xkcd}
comic and subsequently
discussed in blog posts 
\cite{mat_blog, Ilmari_first_links, xkcd}
---we holistically study 
Wikipedia's FLN as a map relating areas of human knowledge. 
The map is a hierarchical taxonomy where, for example, "Train" links to a parent node, 
"Rail Transport," while many child nodes, "Steel" and "Horsepower," link to 
"Train." Unlike previous taxonomies
created by individuals
\cite{locke, descartes, aristotle}
or a select group of individuals 
\cite{hist_thesaurus}, 
the organization of ideas in the FLN 
does not emerge from a concerted or a centralized effort. 
Rather each link in the FLN emerges from the choice of each author.
Collectively, the FLN is a wealth of relations among inventions, places,
figures, objects, and events across space and time.

Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
In a novel approach, we develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of First Links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, indegree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

One essential measure of a directed network's structure is the degree distribution 
\cite{newman}. 
Degree distribution has been used to study many phenomena from disease outbreak 
\cite{disease} 
to the dynamics of social networks 
\cite{social_nets}.
The degree distribution in the FLN describes how many First Links point to a 
particular article. 
We measure the indegree in the FLN to gauge the number of articles referencing a particular article.
A higher indegree indicates more authors selected the particular article
as a relevant reference. On the other hand, articles with zero 
indegree have no references---they are outer leaves in the FLN. 
Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
We can rank articles by indegree to see which articles are most and least referenced in the FLN. 
Additionally, we can assess the degree distribution to see whether authors tend to reference all articles
equally or only a few, as observed in many social networks \cite{social_nets}.

The degree distribution captures one step in the FLN by measuring   
the destination of a given link. 
Indegree measures only the particular set of First Links between two articles, 
rather than the richer dynamics of how many articles fit together---how links 
create a flow through the FLN.
One way to expand the analysis beyond a single step is to map the full path of 
an article's First Link. 
Each article is both a destination and a link to another article, similar to an element 
in a linked list data structure. 
By following the chain of links we can construct a path through the FLN with each article 
leading to the next through the First Link. 
For example "Train" has a First Link to "Conveyance of Passengers and Goods," which is itself
an article with a First Link to "Goods" and so on. The path starting at "Train" 
is the collection of articles obtained by following the First Links starting at "Train"
(see figure~\ref{fig:Train First Links}).
The path contains "Goods," "Economics," "Social Science," and other articles related to 
"Train," but not directly referenced by a First Link.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/traversal_visit_algo_figure.pdf}  
  \caption{
    \textbf{Traversal Visit Algorithm on a sample network.}
     The traversal visit vectors are an adjacency matrix for the paths through the network: the first column is indicates the path formed starting with article A. The number of traversal visits for article A is then the number of paths containing A or the sum of the first row in our matrix:
     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
  }
  \label{fig:Traversal Visits}
\end{figure*}

Applying a similar concept with  
varying starting articles, we form 
paths through the FLN moving from one article to the next.
A path in the FLN connects a group of articles, allowing us to measure how two articles relate,
even if the articles are not directly linked. 
Path-connected articles express a flow of concepts.
The notion of paths allows us to map not only which articles link directly to "Philosophy" 
for example, but also which ideas flow to "Philosophy":
path-connected articles to "Philosophy."
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Starting at each article, we construct a path through the FLN, 
ultimately mapping a flow of connections among articles.

@@

Our goal is to map the flow of First Links by following the paths of First Links 
starting at each article. 
Since a path reveals how the ideas in one article may eventually be 
anchored in another, the first metric we develop quantifies the accumulation of First Links.
The algorithm begins by selecting an article, then traversing the path formed
by following the First Links. Each time a First Link references an article, we increment a count
associated with the article. 
We continue until the First Link is repeated or is invalid---defined here to mean outside of Wikipedia.
We select a second article and repeat the process until we have 
constructed a path for each article in the network. We call the resulting count for each article the number of {\it traversal visits}. The number of traversal visits of an article 
measures the number of references flowing to the ideas in the article.

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/traversal_funnel_algo_figure.pdf}
  \caption{
    \textbf{Traversal Funnel Algorithm on a sample network.}
  The algorithm for traversal funnels is identical to the previous algorithm for traversal visits with one alteration: the path ends at the start of a cycle to distinguish articles directing a path into a cycle from articles that simply happen to be in a highly traversed path. We can construct similar vectors by considering each path through the network, measuring traversal funnels for a particular article as the sum of the entries in its corresponding row. For example
  the number of traversal funnels for article $E$ is 
  $\sum_{i=1}^7 E_{\text{funnel, i}} = 2$.}
  \label{fig:Traversal Funnels}

We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
(see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
first column in the traversal vists matrix. 
An entry of $1$ indicates the path contains a given article and 
$0$ indicates the path doesn't. The columns identify the starting article of the path. 
For example, the second column corresponds to the path starting at article B. 
To compute the number traversal visits for article A we sum the corresponding row in our matrix.
The total is the number of paths containing article A. 
Analogous to the flow 
in a river network, traversal visits measure the accumulation of water at point A. 
By constructing a traversal visits matrix, we can similarly compute the number of traversal 
visits by summing along the corresponding row. The traversal visits matrix for 
Wikipedia consists of 121 million entries to account for each path through the FLN. 
We measure the accumulation of First Link references by summing the corresponding row for the 
article of interest.

In addition to accumulation, the traversal visits algorithm creates paths 
which uncover depth in the FLN. 
By measuring the number of First Links between
two articles, we get an additional piece of information we call 
{\it path length.} We can compute the length of a path by summing 
along a column in our traversal visits matrix. 
The path length describes how closely related ideas are, gauging the FLN's depth.
The greater the number of First Links separating two articles, the greater the number of ideas needed to relate their corresponding ideas. Although
"Train" is related to "Economics" for example, there are several articles
bridging the connection:
"Train" is more specifically related to transportation, whose object
is often goods. Goods are one of the fundamental objects of study in Economics.
Described in links, this relationship is captured by a relatedness of $4$ 
First Links ultimately connecting "Train" to "Economics." 
We can examine path length as a relative measure of how closely connected ideas
may be, but also as an aggregate measure describing the overall separation or 
connectedness in the FLN.
We can characterize whether ideas are closely connected by a handful of links, 
many, or whether a more nuanced organization exists.

One possible path through the FLN is a {\it cycle} or a group of articles 
linking to one another inside a loop. In our sample network 
(see figure~\ref{fig:Traversal Visits}) 
a cycle exists among nodes A, B, and C. 
Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
By recording the history of the articles along a path in the FLN, we can identify
the types of cycle structures within the FLN. Furthermore, since each article 
has an associated traversal visit count, we can rank cycles by 
the number of references directed towards a cycle. 
We can also form and rank {\it basins} in the network by identifying groups of 
path-connected articles, not necessarily forming a perfect cycle.
A basin connects a group of articles and identifies the paths 
to a particular article.
By identifying basins and cycles, we pinpoint groups of closely related articles.

While traversal visits measure accumulation, each article's First Link also 
influences the shape of the FLN. 
At a large point of accumulation, a single article's First Link 
can exert great influence over the shape of the FLN by directing many
references on a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many First Links, we develop a second metric called {\it traversal funnels}.

To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle, from an article exerting influence over where the First Links flow. 
An article with a large number of traversal funnels directs many references
towards a particular cycle. In our sample network 
(see figure~\ref{fig:Traversal Funnels}) article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
recipients of the flow---without exerting any direct influence themselves. 
While articles A, B, and C all have the same number of traversal visits, 
article C has 4 traversal funnels (A and B have none). By incrementing the count of 
First Link references on paths only up to a cycle, we can identifying which articles
exert the greatest influence over the FLN's structure.

By studying the FLN not only as collection of directly linked pairs of articles, but
as a flow, 
we build a powerful arsenal of information with which to study the FLN. 
From accumulated references, cycles and basins to influence, we can measure how the many articles in Wikipedia are organized and related.

@@

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
  \caption{
    \textbf{Highest Ranking Articles by indegree.}
  }
  We rank each article by the number of direct First Links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
  comprised of various facets.
  \label{fig:ndegree list}
\end{figure}

We identify the set of First Links directly referencing a particular article 
by studying the indegree of the FLN. 
A higher indegree suggests more authors found the ideas in the article 
pertinent to reference. 
We rank all 11 million articles by indegree to find 
the "United States" with 80,249 direct First Links as the most referenced
Wikipedia article 
(see figure~\ref{fig:ndegree list}). 
Other high-ranking articles
include foundational abstract concepts such as "village," "species,"; 
sports associations such as "American Football," "Association Football"; 
and developed nations such as "France," "Japan," "Russia," "Australia," and 
the "Netherlands." These high ranking articles are useful abstractions: nations
describe a collection of individuals, a common culture, language, or 
geographical proximity; sports teams describe an ever changing collection of 
sports players often associated with a cultural identity and a geographical 
region. Since abstractions such as nations and teams are inherently comprised
of many parts, authors reference the larger abstraction when describing 
a part. For example, in describing the geographical region of north America, 
the United States, Canada, or Mexico are natural references anchoring the 
geographical location to a larger national identity.

"Philosophy" and other philosophical concepts with many traversal visits
are not among the highest-ranking articles by indegree.
"Philosophy" has an indegree of 581, with direct First Links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy" (see the coming traversal visits discussion), 
the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of First Links, as we argue in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when generalizing from specific articles towards broader notions.

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution.}
  }
  We construct a log-log plot and fit our results with a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears to be scale free.
  \label{fig:degree distribution}
\end{figure}

The FLN's indegree exhibits a scale-free distribution where a few articles 
receive a most direct first references, while most articles receive few or none.
The average indegree for all 11 million articles is 3.6 direct First Links with a standard deviation of 89.5 links.
Only 4,826 articles have more than 100 direct First Links and $75\%$ of articles
have fewer than 9. 
When fit with a linear model on a log-log scale (log(rank) versus log(indegree)), 
the model's r value is $-0.98$ suggesting a strong log-log linear fit 
with a power law exponent of $-0.79$.

@@

Beyond indegree, we develop the notion of a path to describe the structure of the FLN as a 
flow relating ideas. We traverse every possible path through the FLN, taking 
$232$ million steps along the way.

@@

We first seek to describe the depth of the FLN: how many links does a 
connection of ideas span? We gauge the FLN's depth via path length or the number of links traversed until a repeated or invalid link (one outside the FLN). 
We discover the longest path length is 365,
corresponding to the yearly calendar of Orthodox Liturgics.
Each day's Liturgics links to the next day's. On the last calendar day, the last article simply links back to January 1, forming a 365-cycle 
(see discussion of cycles).
We also found similarly lengthy paths following the evolution of a place or topic through time: 
"1953 in Scotland" or "1560s Architecture", with articles sequentially proceeding by year, decade or era.
The longest paths connecting ideas are organized temporally.

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
  \caption{
    \textbf{Path Length Distribution.}
  }
  The median network depth measured by the number of First Links in a path
  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
  other historical articles about "Scotland," "UK" and so on are slightly outside
  the third quartile. More than $75\%$ of articles have path lengths between 
  $0$ and $50$ links.
